---
title: "Reproducible Research Week 2 Project"
author: "Aditya Das"
date: "20 October 2017"
output: html_document
---
## Loading and preprocessing the data

1. Load the data
2. Process the data into a format suitable for analysis

Lets have a look at the summary of the data and five rows of the data
```{r echo=FALSE, results='hide'}
library(ggplot2)
```

```{r}
actidata<- read.csv("activity.csv")
summary(actidata)
head(actidata)
```

## What is mean total number of steps taken per day?

1.Calculate the total number of steps taken per day

```{r}
actidata$date<- as.Date(as.character(actidata$date,"%Y-%m-%d"))
steps<- aggregate(actidata$steps, by = list(Date = actidata$date), FUN = sum)
names(steps)[names(steps) == "x"]<- "Total"
head(steps)
```
2. Make histogram of the number of steps taken per day
```{r}
hist1 <- ggplot(data = na.omit(steps), aes(Total)) + 
  geom_histogram(binwidth = 1500, colour = "white") + 
  xlab("Total Number of Steps Taken Each Day") + 
  ylab("Count") + 
  ggtitle("Histogram of the Total Number of Steps Taken Each Day")
print(hist1)
```

3.Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(na.omit(steps$Total))
median(na.omit(steps$Total))
```

## What is the average daily activity pattern?

1.Make a time series plot (i.e. ???????????????? = "????") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis) type = "l" means the plot is line graph.

```{r}
five_min_steps <- aggregate(steps ~ interval, data = actidata, FUN = mean)
TimeSeries1 <- ggplot(data = five_min_steps, aes(x = interval, y = steps)) + 
    geom_line() +
    xlab("Time Intervals (5 Minutes is an unit)") + 
    ylab("Total Number of Steps") +
    ggtitle("Average Number of Steps Taken of the 5-Minute Interval")
print(TimeSeries1)
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
head(five_min_steps)
five_min_steps[which(five_min_steps$steps == max(five_min_steps$steps)),]
```

## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as ????????). The presence of missing days may introduce bias into some calculations or summaries of the data.

1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sapply(X = actidata, FUN = function(x) sum(is.na(x)))
```

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc. I will use the mean for that 5 -minute interval to replace all the missing values in the dataset. At the end, I will check if all the NAs have been replaced.

```{r}
library(dplyr)
replace_with_mean <- function(num) replace(num, is.na(num), mean(num, na.rm = TRUE))
meanday <- (actidata %>% group_by(interval) %>% mutate(steps = replace_with_mean(steps)))
head(meanday)
sum(is.na(meanday))
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
new_dataset <- as.data.frame(meanday)
head(new_dataset)
summary(new_dataset)
```

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps? We make a histogram of the total number of steps taken each day first by using the new version dataset

```{r}
new_steps <- aggregate(new_dataset$steps, by = list(new_dataset$date), FUN = sum)
names(new_steps)[names(new_steps) == "x"] <- "Total"
names(new_steps)[names(new_steps) == "Group.1"] <- "Date"
hist2 <- ggplot(data = new_steps, aes(Total)) + 
    geom_histogram(binwidth = 1500, colour = "white") +
    xlab("Total Number of Steps Taken Each Day") +
    ylab("Count") +
    ggtitle("Histogram of the Total Number of Steps Taken Each Day with New Version Dataset")
print(hist2)
```

Now we compare two plots


```{r}
library(grid)
library(gridExtra)
grid.arrange(hist1, hist2, ncol = 2)
```


Compare the mean and the median


```{r}
mean(na.omit(steps$Total))
median(na.omit(steps$Total))
mean(new_steps$Total)
median(new_steps$Total)
```

From the comparison, we can see that the highest count of the new version data is larger than the one we have with NAs. The means of each dataset are same. The medians of each dataset are slightly different.

## Are there differences in activity patterns between weekdays and weekends?
Use the dataset with the filled-in missing values which is called new_steps dataset.

1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.


```{r}
new_dataset$WeekendOrWeekday <- ifelse(weekdays(as.Date(new_dataset$date)) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "Weekday", "Weekend")
head(new_dataset)
```


2. Make a panel plot containing a time series plot (i.e. ???????????????? = "????") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
new_dataset <- (new_dataset %>% group_by(interval, WeekendOrWeekday) %>% summarise(Mean = mean(steps)))
ggplot(new_dataset, mapping = aes(x = interval, y = Mean)) + geom_line() +
    facet_grid(WeekendOrWeekday ~.) + xlab("Interval") + ylab("Mean of Steps") +
    ggtitle("Comparison of Average Number of Steps in Each Interval")
```


There are differences in activity patterns between weekdays and weekends. Compared with weekdays, during weekends, people tend to get up late and stay up late. On weekends, people tend to be more active, since their steps are not stable caused of not working.